---
title: "Chapter 7"
author: "Julin N Maloof"
date: "3/20/2018"
output: 
  html_document: 
    keep_md: yes
---

# Chapter 7

6, 7, 8, 9

## Q6

_6. In this exercise, you will further analyze the Wage data set considered throughout this chapter._

_(a) Perform polynomial regression to predict wage using age. Use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the results of hypothesis testing using ANOVA? Make a plot of the resulting polynomial fit to the data._

```{r}
library(ISLR)
library(tidyverse)
library(broom)
library(splines)
data(Wage)
?Wage
head(Wage)
```


```{r}
fits.poly <- tibble(degree=1:15)
fitpoly <- function(degree,data=Wage) {
  lm(wage ~ poly(age,degree=degree), data=data)
}
fits.poly <- fits.poly %>% mutate(fit=map(degree, fitpoly))
fits.poly # a column of models...
```

some stats for each model
```{r}
fits.poly <- fits.poly %>% mutate(glance=map(fit,glance))
fits.poly %>% unnest(glance)
```

```{r}
fits.poly$fit %>% do.call(anova,.)
```

So...anova suggests that degree=3 is best.  This also is consistent with the adjusted R-squared.

get predictions from degree=3 model
```{r}
pred.df <- tibble(age=seq(min(Wage$age),max(Wage$age),by=1))
pred.df <- cbind(pred.df, predict(fits.poly$fit[[3]], newdata=pred.df, se.fit=TRUE))
pred.df <- pred.df %>% mutate(upper.95 = fit + 2*se.fit, lower.95 = fit - 2*se.fit)
```

plot it
```{r}
ggplot(pred.df,aes(x=age,y=fit,ymin=lower.95,ymax=upper.95)) +
  geom_ribbon(fill="gray70") +
  geom_line(color="red",lwd=2) +
  geom_point(aes(y=wage,x=age),shape=1,data=Wage,inherit.aes=FALSE) +
  ylab("wage")
```


_(b) Fit a step function to predict wage using age, and perform cross-validation to choose the optimal number of cuts. Make a plot of the fit obtained._

```{r}
fits.cut <- tibble(cuts=2:10)
fitcut<- function(cuts,data=Wage) {
  lm(wage ~ cut(age,cuts), data=Wage)
}
fits.cut <- fits.cut %>% mutate(fit=map(cuts, fitcut))
fits.cut # a column of models...
```

some stats for each model
```{r}
fits.cut <- fits.cut %>% mutate(glance=map(fit,glance))
fits.cut %>% unnest(glance)
```

cross validation
```{r}
k <- 10
folds <- sample(1:k,nrow(Wage),replace = TRUE)

fitcut.cv <- function(cuts,folds,data=Wage) {
  sapply(1:max(folds), function(fold) {
    data$cut <- cut(data$age,cuts)
    train=data[folds!=fold,]
    test=data[folds==fold,]
    fit <- lm(wage ~ cut, data=train)
    pred.cv <- predict(fit,newdata=test)
    MSE <- (test$wage-pred.cv)^2 %>% mean()
    MSE
  }
  )
}

cv.results <- sapply(2:10,fitcut.cv,folds)

cv.results

cv.summary <- tibble(
  cuts = 2:10,
  cv.mean = colMeans(cv.results),
  cv.se = apply(cv.results,2,sd),
  cv.upper=cv.mean+cv.se,
  cv.lower=cv.mean-cv.se)

cv.summary

cv.summary %>% ggplot(aes(x=cuts,y=cv.mean,ymax=cv.upper,ymin=cv.lower)) +
  geom_ribbon(fill="gray85") +
  geom_point() +
  geom_line() +
  ylim(c(0,2100))
```

Choose 4 cuts; this is where it starts to level off.  Arguably could choose 2 based on the 1SE rule, or 8 based on minimum.

Predict and plot for cuts=4

```{r}
pred.df <- tibble(age=seq(min(Wage$age),max(Wage$age),by=1))
pred.df <- cbind(pred.df, predict(fits.cut$fit[[3]], newdata=pred.df, se.fit=TRUE))
pred.df <- pred.df %>% mutate(upper.95 = fit + 2*se.fit, lower.95 = fit - 2*se.fit)
```

plot it
```{r}
ggplot(pred.df,aes(x=age,y=fit,ymin=lower.95,ymax=upper.95)) +
  geom_ribbon(fill="gray70") +
  geom_line(color="red",lwd=2) +
  geom_point(aes(y=wage,x=age),shape=1,data=Wage,inherit.aes=FALSE) +
  ylab("wage")
```

```{r}
summary(fits.cut$fit[[3]])
```

## Q7

_7. The Wage data set contains a number of other features not explored in this chapter, such as marital status (maritl), job class (jobclass), and others. Explore the relationships between some of these other predictors and wage, and use non-linear fitting techniques in order to fit flexible models to the data. Create plots of the results obtained, and write a summary of your findings._

First make a plot...

```{r}
pairs(Wage)
```

Looks like the potentially interesting variables are maritil, race, and education

```{r}
Wage %>% select(wage,maritl, race, education, health) %>% pairs()
```

OK this doesn't really make since because the rest of these are categorical. but anyway...

### polynomials
```{r}
fit1 <- lm(wage ~ poly(age,3) + poly(as.numeric(maritl),3) + poly(as.numeric(race),3) + poly(as.numeric(education),3) ,data=na.omit(Wage))

summary(fit1)
```

```{r}
fit2  <- lm(wage ~ poly(age,2) + poly(as.numeric(maritl),3) + poly(as.numeric(race),1) + poly(as.numeric(education),2) ,data=Wage)

summary(fit2)
```

cross validation
```{r}

fit.cv <- function(folds,data=Wage) {
  sapply(1:max(folds), function(fold) {
    train=data[folds!=fold,]
    test=data[folds==fold,]
    fit <- lm(wage ~ poly(age,2) + poly(as.numeric(maritl),3) + poly(as.numeric(race),1) + poly(as.numeric(education),2) ,data=train)
    pred.cv <- predict(fit,newdata=test)
    MSE <- (test$wage-pred.cv)^2 %>% mean()
    MSE
  }
  )
}

cv.results <- fit.cv(folds)

cat("MSE: ",mean(cv.results),"\n")

cat("SE of MSE: ",sd(cv.results),"\n")
```

This is a better fit than obtained above.

### splines

```{r}
fit3 <- lm(wage ~ ns(age,5) + ns(as.numeric(maritl),4) + ns(as.numeric(race),3) + ns(as.numeric(education),4) ,data=Wage)

summary(fit3)
```

```{r}
fit4 <- lm(wage ~ ns(age,2) + ns(as.numeric(maritl),2) + ns(as.numeric(race),1) + ns(as.numeric(education),4) ,data=Wage)

summary(fit4)
```

cross validation
```{r}

fit.cv <- function(folds,data=Wage) {
  sapply(1:max(folds), function(fold) {
    train=data[folds!=fold,]
    test=data[folds==fold,]
    fit <- lm(wage ~ ns(age,2) + ns(as.numeric(maritl),2) + ns(as.numeric(race),1) + ns(as.numeric(education),4) ,data=train)
    pred.cv <- predict(fit,newdata=test)
    MSE <- (test$wage-pred.cv)^2 %>% mean()
    MSE
  }
  )
}

cv.results <- fit.cv(folds)

cat("MSE: ",mean(cv.results),"\n")

cat("SE of MSE: ",sd(cv.results),"\n")
```

## Q8

## Q9

_This question uses the variables dis (the weighted mean of distances to five Boston employment centers) and nox (nitrogen oxides concentration in parts per 10 million) from the Boston data. We will treat dis as the predictor and nox as the response._

```{r}
data(Boston,package = "MASS")
qplot(Boston$dis, Boston$nox)
```

_(a) Use the poly() function to fit a cubic polynomial regression to predict nox using dis. Report the regression output, and plot the resulting data and polynomial fits._

```{r}
fit9a <- lm(nox ~ poly(dis,3), data=Boston)
summary(fit9a)
dis.grid <- seq(min(Boston$dis), max(Boston$dis), by=0.1)
pred.9a <- as.data.frame(predict(fit9a, newdata = list(dis=dis.grid),se=TRUE)) %>%
  select(-df, -residual.scale)
pred.9a$dis <- dis.grid
head(pred.9a)
```

```{r}
pl <- ggplot(pred.9a,aes(x=dis,y=fit, ymin=fit-se.fit, ymax=fit+se.fit)) +
  geom_ribbon(fill="gray85") +
  geom_line(color="blue") +
  geom_point(aes(x=dis, y=nox), data=Boston, shape=1, inherit.aes = FALSE)
pl
```


_(b) Plot the polynomial fits for a range of different polynomial degrees (say, from 1 to 10), and report the associated residual sum of squares._

```{r}
fitpoly <- function(degree,data) {
  lm(nox ~ poly(dis,degree=degree), data=data)
}

fits.9b <- tibble(degree=1:10)
fits.9b <- fits.9b %>% mutate(fit=map(degree,fitpoly,Boston))
fits.9b %>% mutate(glance=map(fit,glance)) %>% unnest(glance)
```

```{r, eval=FALSE}



make.plot <- function(fit) {
  
  pred.tmp <- as.data.frame(predict(fit, newdata = list(dis=dis.grid),se=TRUE)) %>%
  select(-df, -residual.scale)
pred.tmp$dis <- dis.grid
  
  ggplot(pred.tmp,aes(x=dis,y=fit, ymin=fit-se.fit, ymax=fit+se.fit)) +
  geom_ribbon(fill="gray85") +
  geom_line(color="blue") +
  geom_point(aes(x=dis, y=nox), data=Boston, shape=1, inherit.aes = FALSE)
}

for(i in 1:10) {
  print(make.plot(fits.9b$fit[[i]]))
}

```


_(c) Perform cross-validation or another approach to select the optimal degree for the polynomial, and explain your results._

_(d) Use the bs() function to fit a regression spline to predict nox using dis. Report the output for the fit using four degrees of freedom. How did you choose the knots? Plot the resulting fit._

_(e) Now fit a regression spline for a range of degrees of freedom, and plot the resulting fits and report the resulting RSS. Describe the results obtained._

_(f) Perform cross-validation or another approach in order to select the best degrees of freedom for a regression spline on this data. Describe your results._

