---
title: "Chapter 6"
author: "Julin N Maloof"
date: "2/27/2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE)
library(ISLR)
library(glmnet)
library(tidyverse)
library(leaps)
library(pls)
```


## Q2
_For parts (a) through (c), indicate which of i. through iv. is correct. Justify your answer._
_i. More flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance._
_ii. More flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias._
_iii. Less flexible and hence will give improved prediction accuracy when its increase in bias is less than its decrease in variance._
_iv. Less flexible and hence will give improved prediction accuracy when its increase in variance is less than its decrease in bias._

### _(a) The lasso, relative to least squares, is:_

iii.  Because lasso forces some coefficients to zero it is less flexible.  This will cause an increase in bias but a decrease in variance.

### _(b) Repeat (a) for ridge regression relative to least squares._

iii.  Because lasso forces some coefficients to be small it is less flexible.  This will cause an increase in bias but a decrease in variance.


### _(c) Repeat (a) for non-linear methods relative to least squares._

ii.  Because the model is not constrained to be linear there is more flexibility.  This flexibility can reduce bias but also will increase variance.

## Q3

_Asks about adding an additional constraint to a standard least squares where we limit the size of the beta coefficients to be less than some value _s_._

_For parts (a) through (e), indicate which of i. through v. is correct. Justify your answer._

_i. Increase initially, and then eventually start decreasing in an inverted U shape._
_ii. Decrease initially, and then eventually start increasing in a U shape._
_iii. Steadily increase._
_iv. Steadily decrease._
_v. Remain constant._

_(a) As we increase s from 0, the training RSS will:_

iv. Steadliy decrease.  We have a more and more flexible model so it will better be able to fit the training set.  Actually it won't steadly decrease, it will plateau...

_(b) Repeat (a) for test RSS._

I think this depends on the scenario.  If we have a small number of good predictors then this will be pretty similar to (a).  OTOH if we have a large number of predictors, or some that aren't good predictors, then 

ii. Decrease initially, and then eventually start increasing in a U shape.

_(c) variance_

Variance is high if there are a lot of differences in models when we use different training sets.

with low s we won't fit anything well (and the model won't change much from training to training) so variance will be low.  As s increases there will start to be more and more difference betweenthe models, so I expect Variance to increase.

_(d) Squared bias._

Bias represents the difference between our model and the data.  When s = 0 then there is no flexibility and bias will be high.  This will decrease as s increases.

_(e) irreducible error._

By definition this should remain constant.

## Q4

_This asks the same questions as 3, but here the model is the Ridge Regression model and penalizes based on lambda, the tuning coefficent that is amultiplier for the sum of the squred beta coefficients._

_As we increase lambda from 0..._

_(a) training RSS_

Will steadily decrease as we fit less and less flexible models

_(b) test RSS_

Will decrease as we reduce variance, and then increase as eventually we increase bias

_(c) variance_

Will decrease steadily

_(d) bias_

will increase steadily

_(e) irreducible error_

unchanged

## Q5 

## Q8

## Q9 

_9. In this exercise, we will predict the number of applications received using the other variables in the College data set._

_(a) Split the data set into a training set and a test set._

```{r}
data(College)
head(College)
College$Private <- as.numeric(College$Private)-1 # 0 = no, 1 = yes
set.seed(123)
train <- sample(c(TRUE,FALSE),size = nrow(College),prob = c(0.80,0.20),replace=TRUE)
college.train <- College %>% as_tibble() %>% filter(train)
college.test <- College %>% as_tibble() %>% filter(!train)
```


_(b) Fit a linear model using least squares on the training set, and report the test error obtained._

```{r}
lm9b <- lm(Apps ~ . , data = college.train)
summary(lm9b)
```

```{r}
lm9c.predict <- predict(lm9b,newdata = college.test)
test.error <- (college.test$Apps - lm9c.predict)^2 %>% mean()
test.error
```


_(c) Fit a ridge regression model on the training set, with λ chosen by cross-validation. Report the test error obtained._

```{r}
ridge9c <- college.train %>% select(-Apps) %>% as.matrix() %>%
  glmnet(y=college.train$Apps, alpha = 0)
```

```{r}
cv.out <- college.train %>% select(-Apps) %>% as.matrix() %>%
  cv.glmnet(y=college.train$Apps, alpha = 0)
```

```{r}
plot(cv.out)
lam1se=cv.out$lambda.1se
lam1se
log(lam1se)
```

```{r}
ridge9c.predict <- college.test %>% select(-Apps) %>% as.matrix() %>%
  predict(ridge9c,s=lam1se,newx=.)
(college.test$Apps - ridge9c.predict)^2 %>% mean()
```
Something wrong??

```{r}
ridge9c.predict <- college.test %>% select(-Apps) %>% as.matrix() %>%
  predict(ridge9c,s=cv.out$lambda.min,newx=.)
(college.test$Apps - ridge9c.predict)^2 %>% mean()
```

Still high...

_(d) Fit a lasso model on the training set, with λ chosen by cross- validation. Report the test error obtained, along with the number of non-zero coefficient estimates._

```{r}
lasso9d <- college.train %>% select(-Apps) %>% as.matrix() %>%
  glmnet(y=college.train$Apps, alpha = 1)
```

```{r}
cv.out <- college.train %>% select(-Apps) %>% as.matrix() %>%
  cv.glmnet(y=college.train$Apps, alpha = 1)
```

```{r}
plot(cv.out)
lam1se=cv.out$lambda.1se
lam1se
log(lam1se)
```

```{r}
lasso9d.predict <- college.test %>% select(-Apps) %>% as.matrix() %>%
  predict(lasso9d,s=lam1se,newx=.)
(college.test$Apps - lasso9d.predict)^2 %>% mean()
```
Larger than lm.  Something wrong??

```{r}
lasso9d.predict <- college.test %>% select(-Apps) %>% as.matrix() %>%
  predict(lasso9d,s=cv.out$lambda.min,newx=.)
(college.test$Apps - lasso9d.predict)^2 %>% mean()
```

```{r}
lasso9d.predict <- college.test %>% select(-Apps) %>% as.matrix() %>%
  predict(lasso9d,s=0,newx=.)
(college.test$Apps - lasso9d.predict)^2 %>% mean()
```

## Q11
_We will now try to predict per capita crime rate in the Boston data set._

### a
_Try out some of the regression methods explored in this chapter, such as best subset selection, the lasso, ridge regression, and PCR. Present and discuss results for the approaches that you consider._

Prepare the data
```{r}
data(Boston, package = "MASS")
boston <- as_tibble(Boston)
boston
```

#### Best Subset

```{r}
bs <- regsubsets(crim ~ .,data=boston,nvmax = 13)
(bs.sum <- summary(bs))
plot(bs)
```

make my own plot
```{r}
bs.plot.frame <- tibble(predictors=1:13,bs.sum$rsq, bs.sum$adjr2, bs.sum$cp, bs.sum$bic) %>%
  gather(key="criterion",value="value",-predictors)
bs.plot.frame %>% ggplot(aes(x=predictors,y=value)) +
  geom_line() +
  facet_wrap(~criterion,scales="free_y")
```

Get pretty different answers here.  Try a CV approach.


```{r}
#From ISLR book
predict.regsubsets  <- function (object ,newdata ,id ,...){
 form=as.formula(object$call [[2]])
 mat=model.matrix(form,newdata)
 coefi=coef(object ,id=id)
 xvars=names(coefi)
 mat[,xvars]%*%coefi}
```

create folds
```{r}
k <- 10
set.seed(123)
folds <- sample(1:k,nrow(boston),replace=TRUE)
cv.errors <- matrix(NA,k,13, dimnames=list(NULL, paste(1:13)))
```

test the folds
```{r}
for(j in 1:k){
  best.fit <- regsubsets(crim ~ . , data=boston[folds!=j,],nvmax = 13)
  for(i in 1:13) {
    pred=predict(best.fit, boston[folds==j,],id=i)
    cv.errors[j,i] <- mean( (boston$crim[folds==j]-pred)^2)
  } 
}
```

check the CV MSE
```{r}
(mean.cv.errors <- apply(cv.errors ,2,mean))
plot(mean.cv.errors, type='b')
```
12 predictors in best... refit model
```{r}
reg.best <- regsubsets(crim ~ ., data=boston , nvmax=12)
coef(reg.best,12)
```

#### ridge

```{r}
mod.matrix <- model.matrix(crim ~ . -1., data = boston)
boston.ridge <- glmnet(mod.matrix, boston$crim, alpha=0)
boston.ridge.cv <- cv.glmnet(mod.matrix, boston$crim, alpha=0, foldid=folds)
plot(boston.ridge.cv)
```

```{r}
boston.ridge.cv$lambda.min
boston.ridge.cv$lambda.1se
boston.ridge.cv$cvm[boston.ridge.cv$lambda==boston.ridge.cv$lambda.min]
boston.ridge.cv$cvm[boston.ridge.cv$lambda==boston.ridge.cv$lambda.1se]
coef(boston.ridge,s=boston.ridge.cv$lambda.min)
coef(boston.ridge,s=boston.ridge.cv$lambda.1se)
```

#### lasso

```{r}
mod.matrix <- model.matrix(crim ~ . -1., data = boston)
boston.lasso <- glmnet(mod.matrix, boston$crim, alpha=1)
boston.lasso.cv <- cv.glmnet(mod.matrix, boston$crim, alpha=1, foldid=folds)
plot(boston.lasso.cv)
```

```{r}
boston.lasso.cv$lambda.min
boston.lasso.cv$lambda.1se
boston.lasso.cv$cvm[boston.lasso.cv$lambda==boston.lasso.cv$lambda.min]
boston.lasso.cv$cvm[boston.lasso.cv$lambda==boston.lasso.cv$lambda.1se]
coef(boston.lasso,s=boston.lasso.cv$lambda.min)
coef(boston.lasso,s=boston.lasso.cv$lambda.1se)
```
#### PCR

```{r}
set.seed(1)
boston.pcr <- pcr(crim ~ . , data=boston, scale=TRUE, validation="CV")
summary(boston.pcr)
```

```{r}
validationplot(boston.pcr, val.type = "MSEP")
```

So something like 3 components is good

#### PLS

```{r}
set.seed(1)
boston.pls <- plsr(crim ~ . , data=boston, scale=TRUE, validation="CV")
summary(boston.pls)
```

```{r}
validationplot(boston.pls, val.type = "MSEP")
```


### b
_Propose a model (or set of models) that seem to perform well on this data set, and justify your answer. Make sure that you are evaluating model performance using validation set error, cross- validation, or some other reasonable alternative, as opposed to using training error._

### c
_Does your chosen model involve all of the features in the data set? Why or why not?_
