---
title: "Chapter 3"
author: "Julin N Maloof"
date: "12/5/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dec 6th

Questions 8, 13 (13g optional), 15(a)

## Q1


_Describe the null hypotheses to which the p-values given in Table 3.4 correspond. Explain what conclusions you can draw based on these p-values. Your explanation should be phrased in terms of sales, TV, radio, and newspaper, rather than in terms of the coefficients of the linear model._

* __Intercept:__ the null hypothesis is that there are no sales when there is no advertising.  The p-value is significant, indicating that there are some sales without advertising.
* __TV:__ the null hypothesis is that changing TV advertising while holding radio and newspaper constant does not change sales. The significant p-value indicates that changing TV advertising does affect sales when holding the other predictors constant.
* __radio:__ the null hypothesis is that changing radio advertising while holding TV and newspaper constant does not change sales. The significant p-value indicates that changing radio advertising does affect sales when holding the other predictors constant.
* __newspaper:__ the null hypothesis is that changing newspaper advertising while holding radio and TV constant does not change sales. The insignificant p-value indicates that changing newspaper advertising does not affect sales when holding the other predictors constant.


## Q8

_8. This question involves the use of simple linear regression on the Auto data set._

```{r}
library(ISLR)
library(tidyverse)
data(Auto)
auto <- as.tibble(Auto)
auto
```

_(a) Use the lm() function to perform a simple linear regression with mpg as the response and horsepower as the predictor. Use the summary() function to print the results. Comment on the output. For example:_

```{r}
lm1 <- lm(mpg ~ horsepower, data = auto)

summary(lm1)
```


_i. Is there a relationship between the predictor and the response?_

Yes, there is a highly significant (negative) relationship between horsepower and mpg

_ii. How strong is the relationship between the predictor and the response?_

given very low p-value and very non-zero t-value,  strong

_iii. Is the relationship between the predictor and the response positive or negative?_

negative

_iv. What is the predicted mpg associated with a horsepower of 98? What are the associated 95 % confidence and prediction intervals?_

```{r}
cat("predition at horsepower 98 and 95% confidence intervals:\n")
predict(lm1,data.frame(horsepower=98),interval = "conf")
cat("\npredition at horsepower 98 and 95% prediction intervals:\n")
predict(lm1,data.frame(horsepower=98),interval = "pred")
```

_(b) Plot the response and the predictor. Use the abline() function to display the least squares regression line._

```{r}
auto %>% 
  ggplot(aes(x=horsepower, y=mpg)) +
           geom_point() +
           geom_smooth(method = "lm", se=FALSE)

```



_(c) Use the plot() function to produce diagnostic plots of the least squares regression fit. Comment on any problems you see with the fit._

```{r}
library(modelr)
auto %>% 
  add_residuals(lm1) %>%
  ggplot(aes(x=horsepower, y = resid)) +
  geom_point() +
  geom_hline(yintercept=0)
```
Residuals not evenly distributed!


```{r}
plot(lm1,ask = FALSE)
```

## Q 13

_13. In this exercise you will create some simulated data and will fit simple linear regression models to it. Make sure to use set.seed(1) prior to starting part (a) to ensure consistent results._

_(a) Using the rnorm() function, create a vector, x, containing 100 observations drawn from a N (0, 1) distribution. This represents a feature, X._
_(b) Using the rnorm() function, create a vector, eps, containing 100 observations drawn from a N(0,0.25) distribution i.e. a normal distribution with mean zero and variance 0.25._
_(c) Using x and eps, generate a vector y according to the model Y =−1+0.5X+ε. (3.39)_
_What is the length of the vector y? What are the values of β0 and β1 in this linear model?_

```{r}
set.seed(1)
data13 <- tibble(
  x = rnorm(100,mean = 0, sd = 1),
  eps = rnorm(100,0, sqrt(0.25)),
  y = 1 + 0.5*x + eps)

data13
  
```

$$\beta_0 = 1$$

$$\beta_1=0.5$$
length of y is 100

_(d) Create a scatterplot displaying the relationship between x and y. Comment on what you observe._

```{r}
data13 %>% qplot(x,y,data=.)
```

Pretty nice linear relationship, but some scatter...

_(e) Fit a least squares linear model to predict y using x. Comment on the model obtained. How do βˆ0 and βˆ1 compare to β0 and β1?_

```{r}
lm13 <- lm(y~x, data=data13)
summary(lm13)
confint(lm13)
```

The estimates are near the real values and the 95% confidence intervals include the real values

_(f) Display the least squares line on the scatterplot obtained in (d). Draw the population regression line on the plot, in a different color. Use the legend() command to create an appropriate legend._

```{r}
data13 %>% ggplot(aes(x=x,y=y)) +
  geom_point() +
  geom_smooth(aes(color="blue"),method="lm",se=FALSE) +
  geom_abline(aes(intercept=1,slope=.5,color="red")) +
  scale_color_manual(name="legend",values=c("red"="red","blue"="blue"),labels=c("population","sample"), guide="legend")
```

(g) Now fit a polynomial regression model that predicts y using x and x2. Is there evidence that the quadratic term improves the model fit? Explain your answer.

```{r}
lm13b <- lm(y ~ x + I(x^2), data=data13)
summary(lm13b)
confint(lm13b)
```

The fits are pretty similar, so no real improvement.  RSE and R-squeared similar.

_(h) Repeat (a)–(f) after modifying the data generation process in such a way that there is less noise in the data. The model (3.39) should remain the same. You can do this by decreasing the variance of the normal distribution used to generate the error term ε in (b). Describe your results._

```{r}
data13h <- tibble(
  x = rnorm(100,mean = 0, sd = 1),
  eps = rnorm(100,0, sqrt(0.1)),
  y = 1 + 0.5*x + eps)

data13h
```

```{r}
lm13h <- lm(y ~ x, data=data13h)
summary(lm13h)
confint(lm13h)
```

```{r}
data13h %>%
  ggplot(aes(x=x,y=y)) +
  geom_point() +
  geom_smooth(aes(color="blue"),method="lm",se=FALSE) +
  geom_abline(aes(intercept=1,slope=.5,color="red")) +
  scale_color_manual(name="legend",values=c("red"="red","blue"="blue"),labels=c("population","sample"), guide="legend")
```

smaller confidence intervals, lower RSE, higher R^2

(i) Repeat (a)–(f) after modifying the data generation process in such a way that there is more noise in the data. The model (3.39) should remain the same. You can do this by increasing the variance of the normal distribution used to generate the error term ε in (b). Describe your results.

```{r}
data13i <- tibble(
  x = rnorm(100,mean = 0, sd = 1),
  eps = rnorm(100,0, sqrt(0.5)),
  y = 1 + 0.5*x + eps)

data13h
```

```{r}
lm13i <- lm(y ~ x, data=data13i)
summary(lm13i)
confint(lm13i)
```

```{r}
data13i %>%
  ggplot(aes(x=x,y=y)) +
  geom_point() +
  geom_smooth(aes(color="blue"),method="lm",se=FALSE) +
  geom_abline(aes(intercept=1,slope=.5,color="red")) +
  scale_color_manual(name="legend",values=c("red"="red","blue"="blue"),labels=c("population","sample"), guide="legend")
```

_(j) What are the confidence intervals for β0 and β1 based on the original data set, the noisier data set, and the less noisy data set? Comment on your results._

Already discussed above in part, but...

```{r}
map(list(original=lm13,lower_var=lm13h,higher_var=lm13i),confint)

map(list(original=lm13,lower_var=lm13h,higher_var=lm13i),confint) %>%
  map(as.data.frame) %>%
  map(rownames_to_column,var="coefficient") %>%
  bind_rows(.id="data_source") %>%
  ggplot(aes(x=data_source,ymin=`2.5 %`,ymax=`97.5 %`,color=coefficient)) +
  geom_linerange(lwd=3) + 
  ggtitle("95% Confidence Intervals") +
  geom_hline(yintercept = c(0.5,1))
```

## Q9. _This question involves the use of multiple linear regression on the Auto data set._

_(a) Produce a scatterplot matrix which includes all of the variables in the data set._

```{r}
data(Auto)
auto <- as_tibble(Auto)
pairs(auto)
```

_(b) Compute the matrix of correlations between the variables using the function cor(). You will need to exclude the name variable, which is qualitative._

```{r}
auto %>% select(-name) %>% cor() %>% knitr::kable(digits = 3)
```



_(c) Use the lm() function to perform a multiple linear regression with mpg as the response and all other variables except name as the predictors. Use the summary() function to print the results. Comment on the output._

```{r}
lm.auto <- lm(mpg ~ cylinders +
                displacement +
                horsepower +
                weight +
                acceleration +
                year + 
                origin,
              data=auto)
summary(lm.auto)
```


For instance:
_i. Is there a relationship between the predictors and the re- sponse?_

Yes, based on over-all p-value of < 2.2 e -16

_ii. Which predictors appear to have a statistically significant relationship to the response?_

displacement, weight, year, origin

_iii. What does the coefficient for the year variable suggest?_

mpg increaeses ~ 0.75 per year when other predictors are held constant.

(d) Use the plot() function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers? Does the leverage plot identify any observations with unusually high leverage?

```{r}
plot(lm.auto)
```

## Q10 _This question should be answered using the Carseats data set._

_(a) Fit a multiple regression model to predict Sales using Price,
Urban, and US._

```{r}
data("Carseats")
lm10a <- lm(Sales ~ Price + Urban + US,data = Carseats)
summary(lm10a)
```

_(b) Provide an interpretation of each coefficient in the model. Be
careful—some of the variables in the model are qualitative!_

* Intercept: The predicted base level of saled if carseats were free
* Price: Sales decreased by 54 units per increase in price (if other predicotrs held constant).
* Urban: Are sales affected by urban vs rural store location?  N.S.
* US: Being in the US increases sales by 1200 units (when other factors are constant).

_(c) Write out the model in equation form, being careful to handle the qualitative variables properly._

$$ Sales \sim \beta_0 + \beta_1*Price + \beta_2*Urban + \beta_3*US + \epsilon $$

Or did they mean

$$
Sales = 13.04 - 0.05 * Price - 0.02*UrbanYes + 1.2*USYes
$$
_(d) For which of the predictors can you reject the null hypothesis H0 :βj =0?_

Intercept, price, US

(e) On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.

```{r}
lm10e <- update(lm10a, ~ . - Urban)
summary(lm10e)
```

_(f) How well do the models in (a) and (e) fit the data?_

Adjusted R-squared is a bit below 0.24 meaning there is a lot of unexplained variance.

(g) Using the model from (e), obtain 95% confidence intervals for the coefficient(s).

```{r}
confint(lm10e)
```

(h) Is there evidence of outliers or high leverage observations in the model from (e)?

```{r}
plot(lm10e)
```

nope, looks pretty good.

Also try some 3D plotting:

```{r}
library(plot3D)
library(broom)
library(modelr)
```




## 15a

_15. This problem involves the Boston data set, which we saw in the lab for this chapter. We will now try to predict per capita crime rate using the other variables in this data set. In other words, per capita crime rate is the response, and the other variables are the predictors._

_(a) For each predictor, fit a simple linear regression model to predict the response. Describe your results. In which of the models is there a statistically significant association between the predictor and the response? Create some plots to back up your assertions._

```{r}
library(MASS)
data(Boston)
boston <- as_tibble(Boston)
boston
```

```{r}
predictors <- colnames(boston)[-1]
lmfits <- map(predictors,function(x) lm(crim ~ get(x), data=boston))
```

```{r}
lmsummaries <- lapply(lmfits,summary)
names(lmsummaries) <- predictors
lmsummaries
```

```{r}
sapply(lmsummaries, function(x) x$coefficients["get(x)","Pr(>|t|)"])
```

```{r}
sapply(lmsummaries, function(x) x$adj.r.squared)
```

### OR

(see http://r4ds.had.co.nz/many-models.html)

create a nested data object, one data frame for each predictor
```{r}
library(modelr)
boston.nest <- boston %>% 
  gather(key="predictor",value="value",-crim) %>%
  group_by(predictor) %>%
  nest()
boston.nest # a set of data frames, one for each predictor
boston.nest$data[[1]]
```

Fit a model to each dataframe
```{r}
fitModel <- function(df) lm(crim ~ value, data=df)
boston.nest <- boston.nest%>%
  mutate(model=map(data,fitModel),
         model.summary=map(model,summary))
boston.nest
```

Now, use broom::glance to get summaries
```{r}
library(broom)
boston_glance <- boston.nest %>%
  mutate(glance=map(model,glance)) %>%
  unnest(glance,.drop = TRUE)
boston_glance
```

### old way...

print out the summaries (but no names!)

```{r}
boston.nest$model.summary
```

We can add columns for whatever we want to extract from the summaries....
```{r}
boston.nest <- boston.nest %>%
  mutate(p_value=map_dbl(model.summary,function(x) x$coefficients["value","Pr(>|t|)"]),
         adj.r.squared=map_dbl(model.summary, function(x) x$adj.r.squared))
boston.nest
```

(b) Fit a multiple regression model to predict the response using all of the predictors. Describe your results. For which predictors can we reject the null hypothesis H0 : βj = 0?

```{r}
library(stringr)
fo <- as.formula(str_c("crim ~ ", str_c(predictors, collapse = " + ")))
lm15b <- lm(fo,data = boston)
summary(lm15b)
```

(c) How do your results from (a) compare to your results from (b)? Create a plot displaying the univariate regression coefficients from (a) on the x-axis, and the multiple regression coefficients from (b) on the y-axis. That is, each predictor is displayed as a single point in the plot. Its coefficient in a simple linear regres- sion model is shown on the x-axis, and its coefficient estimate in the multiple linear regression model is shown on the y-axis.
(d) Is there evidence of non-linear association between any of the predictors and the response? To answer this question, for each predictor X, fit a model of the form
Y = β0 +β1X +β2X2 +β3X3 +ε.

